The following is a complete list of articles written about Garnet, as of
September 4, 1994.  There are 40 refereed journal articles, 4 book chapters,
17 tech reports, and 3 articles submitted for publication.


%\vspace{1 line}
{\bf Refereed Journal Articles}

Brad A. Myers. ``User Interface Software Tools.''
{\it ACM Transactions on Computer-Human Interaction}.  To appear.

Brad Vander Zanden, Brad A. Myers, Dario Giuse and Pedro Szekely.
``Integrating Pointer Variables into One-Way Constraint Models,''
{\it ACM Transactions on Computer-Human Interaction}.  Vol. 1, no. 2,
To appear.

Brad A. Myers.  ``Challenges of HCI Design and Implementation,''
{\it ACM Interactions}.  Vol. 1, no. 1.  January, 1994.  pp. 73-83.

Robert J.K. Jacob, John J. Leggett, Brad A. Myers, and Randy Pausch.
``Interaction Styles and Input/Output Devices,'' {\it Behaviour and
Information Technology}.  March-April, 1993. Vol. 12, no. 2.  pp. 69-79.

Dan R. Olsen, Jr., James D. Foley, Scott E. Hudson, James Miller, and
Brad Myers.  ``Research Directions for User Interface Software Tools,''
{\it Behaviour and Information Technology}.  March-April, 1993. Vol. 12,
no. 2. pp. 80-97.

Brad A. Myers and Brad Vander Zanden.  ``Environment for Rapid
Creation of Interactive Design Tools,'' {\it The Visual Computer;
International Journal of Computer Graphics}. Vol 8, No. 2, February,
1992.  pp. 94-116.

Brad A. Myers, Dario A. Giuse, Roger B. Dannenberg, Brad Vander
Zanden, David S. Kosbie, Ed Pervin, Andrew Mickish, and Philippe
Marchal.  ``Garnet; Comprehensive Support for Graphical,
Highly-Interactive User Interfaces,'' {\it IEEE Computer}. Vol. 23, No.
11. November, 1990. pp. 71-85.  Translated into Japanese and reprinted
in {\it Nikkei Electronics}, No. 522, March 18, 1991, pp. 187-205.

Brad A. Myers. ``A New Model for Handling Input,'' {\it ACM Transactions
on Information Systems}.  Vol. 8, No. 3. July, 1990.  pp. 289-320.

Brad Vander Zanden, Brad A. Myers, Dario Giuse and Pedro Szekely.
``Integrating Pointer Variables into One-Way Constraint Models,''
{\it ACM Transactions on Computer-Human Interaction}.  Vol. 1, no.
2, June, 1994. pp. 161-213.



%\vspace{1 line}
{\bf Refereed Conference Articles}

Brad A. Myers. ``The Garnet User Interface Development Environment:
Demonstration Abstract.'' {\it CHI'94 Conference Companion}.  Boston,
MA, Apr. 24-28, 1994. pp. 25-26.

Brad A. Myers and Dan R. Olsen, Jr. ``User Interface Tools: Tutorial
Description'' {\it CHI'94 Conference Companion}.  Boston,
MA, Apr. 24-28, 1994. pp. 421-422.

Brad A. Myers, Dario Giuse, Andrew Mickish, Brad Vander Zanden, David
Kosbie, Richard McDaniel, James Landay, Matthew Goldberg, and Rajan
Parthasarathy. ``The Garnet User Interface Development Environment:
Video Abstract,'' {\it CHI'94 Conference Companion}.  Boston, MA, Apr.
24-28, 1994. pp. 455-456.

Gurminder Singh, Mark Linton, Brad A. Myers, and Marti Szczur. ``From
Research Prototypes to Usable, Useful Systems: Lessons Learned in the
Trenches,'' {\it Proceedings ACM Symposium on User Interface Software
and Technology: UIST'93}. Atlanta, GA, Nov 3-5, 1993.  pp. 139-143.

Brad A. Myers, Richard G. McDaniel, and David S. Kosbie.  ``Marquise:
Creating Complete User Interfaces by Demonstration,'' {\it Proceedings
INTERCHI'93: Human Factors in Computing Systems}.  Amsterdam, The Netherlands,
April 24-29, 1993.  pp. 293-300.

Brad A. Myers, Richard Wolf, Kathy Potosnak, and Chris Graham.
``Heuristics in Real User Interfaces,''  {\it Proceedings
INTERCHI'93: Human Factors in Computing Systems}.  Amsterdam, The
Netherlands, April 24-29, 1993.  pp. 304-307.

James A. Landay and Brad A. Myers. ``Extending an Existing User Interface
Toolkit to Support Gesture Recognition.'' {\it Adjunct Proceedings of
INTERCHI'93}. Amsterdam, The Netherlands, April 24-29, 1993.  pp. 91-92.

Brad A. Myers. ``The Garnet Gilt Interface Builder: Graphical Styles
and Tabs and Techniques for Reducing Call-Back Procedures,''
Application Builder Session, {\it Seventh Annual X Technical
Conference}, Boston, Massachusetts, January 18, 1993.

Osamu Hashimoto and Brad A. Myers. ``Graphical Styles For Building
User Interfaces by Demonstration,'' {\it ACM Symposium on User Interface
Software and Technology}, Monterey, CA, Nov. 16-18, 1992. pp. 117-124.

Brad A. Myers, Dario A. Giuse, and Brad Vander Zanden. ``Declarative
Programming in a Prototype-Instance System: Object-Oriented
Programming Without Writing Methods,'' {\it Proceedings OOPSLA'92: ACM
Conference on Object-Oriented Programming Systems, Languages, and
Applications}.  October 18-22, 1992.  Vancouver, BC, Canada.
{\it SIGPLAN Notices,} vol. 27, no. 10. pp. 184-200.

Brad A. Myers and Mary Beth Rosson.  ``Survey on User Interface
Programming,'' {\it Proceedings SIGCHI'92: Human Factors in Computing
Systems}.  Monterey, CA, May 3-7, 1992.  195-202.

Brad A. Myers. ``Separating Application Code from Toolkits:
Eliminating the Spaghetti of Call-Backs,'' {\it ACM Symposium on User
Interface Software and Technology}, Hilton Head, SC, Nov. 11-13, 1991.
pp. 211-220.

Brad Vander Zanden, Brad A. Myers, Dario Giuse and Pedro Szekely.
``The Importance of Indirect References in Constraint Models,'' {\it ACM
Symposium on User Interface Software and Technology}, Hilton Head, SC,
Nov. 11-13, 1991. pp. 155-164.

Brad A. Myers.  ``Graphical Techniques in a Spreadsheet for Specifying
User Interfaces,'' {\it Proceedings SIGCHI'91: Human Factors in
Computing Systems}.  New Orleans, LA.  April 28-May 2, 1991.  pp.
243-249.

Brad Vander Zanden and Brad A. Myers, ``Automatic, Look-and-Feel
Independent Dialog Creation for Graphical User Interfaces,''
{\it Proceedings SIGCHI '90: Human Factors in Computing Systems}.
Seattle, WA, April 1-5, 1990. pp. 27-34.

Brad A. Myers.  ``Making it Easy to Create Highly-Interactive,
Graphical Applications in Lisp,'' {\it XNextEvent: The Official
Newsletter of XUG, The X User's Group}.  Vol. 3., No. 1. April, 1990.

Brad A. Myers. ``An Object-Oriented, Constraint-Based, User Interface
Development Environment for X in CommonLisp,'' {\it Fourth Annual X
Technical Conference}, Boston, Massachusetts, January 15-17, 1990.

Dario Giuse. ``Efficient Frame Systems,'' {\it Lecture Notes in
Artificial Intelligence, EPIA 1989}.  Vol. 390. J.P. Martins and E.M.
Morgado, eds.  Springer-Verlag, Sep, 1989.

Brad A. Myers, Brad Vander Zanden, and Roger B. Dannenberg. ``Creating
Graphical Interactive Application Objects by Demonstration,'' {\it ACM
Symposium on User Interface Software and Technology}, Williamsburg,
VA, Nov. 13-15, 1989. pp. 95-104.

Charles Wiecha, Stephen Boies, Mark Green, Scott Hudson, and Brad
Myers.  ``Direct Manipulation or Programming: How Should We Design
Interfaces?''  {\it ACM Symposium on User Interface Software and
Technology}, Williamsburg, VA, Nov. 13-15, 1989. pp. 124-126.

Brad A. Myers. ``AI In Demonstrational User Interfaces,'' {\it A New
Generation of Intelligent Interfaces: IJCAI-89 Workshop}, Detroit, MI.
August 22, 1989, pp. 84-91.

Dario Giuse.  ``Efficient Knowledge Representation Systems,''
{\it The Knowledge Engineering Review}, Vol 4, no. 4, 1990.

Brad A. Myers. ``Encapsulating Interactive Behaviors,''
{\it Proceedings SIGCHI '89: Human Factors in Computing Systems}.
Austin, Texas, April 30 - May 4, 1989, pp. 319-324.

Brad A. Myers. ``User Interface Tools: Introduction
and Survey,'' {\it IEEE Software}, Vol. 6, no. 1, Jan, 1989. pp. 15-23. To
be reprinted in {\it Milestones in Software Evolution}, Paul W. Oman and
Ted G. Lewis, ed.  IEEE Computer Society Press.

Pedro Szekely and Brad Myers. ``A User Interface Toolkit Based on
Graphical Objects and Constraints,'' {\it OOPSLA '88: Conference on
Object-Oriented Programming: Systems, Languages and Applications},
San Diego, CA, September 25-30, 1988.
{\it Sigplan Notices}, Vol 23, no. 11, November, 1988. pp. 36 - 45.

Daniel Kuokka and Dario Giuse. ``The Dante Application Interface,''
{\it Proceedings of the 2nd IEEE International Conference on Computer
Workstations}, Clara, California, March 7-10, 1988.

Dario Giuse.  ``Lisp as a Rapid Prototyping Environment: The Chinese Tutor.''
In {\it Lisp and Symbolic Computation - An International Journal},
Kluwer Academic Publishers, May 1987.


%\vspace{1 line}
{\bf Refereed Published Videotapes}

Brad A. Myers, Dario Giuse, Andrew Mickish, Brad Vander Zanden, David
Kosbie, Richard McDaniel, James Landay, Matthew Goldberg, and Rajan
Parthasarathy. {\it The Garnet User Interface Development Environment}.
Technical Video Program of the CHI'94 conference.
{\it SIGGRAPH Video Review}, Issue 97, no. 13.

Brad A. Myers, Andrew Mickish and Osamu Hashimoto.  ``The Garnet Gilt
Interface Builder: Graphical Styles and Tabs and Techniques for
Reducing Call-Back Procedures,'' Application Builder Video Session,
{\it Seventh Annual X Technical Conference}, Boston, Massachusetts,
January 18, 1993.  10 minutes.

Brad Vander Zanden and Brad A. Myers. {\it Creating Graphical
Interactive Application Objects by Demonstration: The Lapidary
Interface Design Tool.} 12 minute videotape.  Technical Video Program
of the SIGCHI'91 conference, New Orleans, LA.  April 28-May 2, 1991.
{\it SIGGRAPH Video Review}, Issue 64, no. 1.

Brad A. Myers. {\it Some of the Widgets}.  17 minute videotape.
Technical Video Program of Interact'90.  Cambridge, England. August
27-31, 1990.

Brad A. Myers. {\it All the Widgets}.  2 hour videotape.  Technical
Video Program of the SIGCHI'90 conference, Seattle, WA.  April 1-4,
1990.  {\it SIGGRAPH Video Review}, Issue 57.



%\vspace{1 line}
{\bf Chapters in Books}


Brad A. Myers. ``User Interface Software,'' {\it Encyclopedia of Computer
Science and Technology}.  Allen Kent and James G. Williams, editors.
New York: Marcel Dekker, Inc., 1994. To appear.

Brad A. Myers. ``State of the Art in User Interface Software Tools,''
{\it Advances in Human-Computer Interaction, Volume 4}.
Edited by H. Rex Hartson and Deborah Hix.  Norwood, NJ: Ablex
Publishing, 1993.  pp. 110-150.

Brad A. Myers.  ``Garnet: Uses of Demonstrational Techniques,''
{\it Watch What I Do: Programming by Demonstration}, Allen Cypher, et. al., eds.
Cambridge, MA: The MIT Press, 1993.  pp. 219-236.

Brad A. Myers.  ``Ideas from Garnet for Future User Interface
Programming Languages,''
{\it Languages for Developing User Interfaces}. Boston: Jones and Bartlett,
1992.  pp. 147-157.


%\vspace{1 line}
{\bf Technical Reports:}


Brad A. Myers.  {\it User Interface Software Tools}. Carnegie Mellon University
School of Computer Science Technical Report, no. CMU-CS-94-182 and
Human Computer Interaction Institute Technical Report CMU-HCII-94-107.
August 1994.

James A. Landay and Brad A. Myers.  {\it Interactive Sketching for the
Early Stages of User Interface Design}.  Carnegie Mellon University
School of Computer Science Technical Report, no. CMU-CS-94-176 and
Human Computer Interaction Institute Technical Report CMU-HCII-94-104.
July 1994.

Brad A. Myers, Dario A. Giuse, Andrew Mickish, and David S. Kosbie.
{\it Making Structured Graphics and Constraints Practical for
Large-Scale Applications}.  Carnegie Mellon University School of
Computer Science Technical Report, no. CMU-CS-94-109 and Human
Computer Interaction Institute Technical Report CMU-HCII-94-100.  May
1994.

Brad A. Myers.  {\it Why are Human-Computer Interfaces
Difficult to Design and Implement?}  Carnegie Mellon University School
of Computer Science Technical Report, no. CMU-CS-93-183.  July 1993.

Brad A. Myers, editor.  {\it The Second Garnet Compendium:
Collected Papers, 1990-1992}.  Carnegie Mellon University School of
Computer Science Technical Report, no. CMU-CS-93-108, February, 1993.

Brad A. Myers, Dario Giuse, Andrew Mickish, Brad Vander Zanden, David
Kosbie, James A. Landay, Richard McDaniel, Rajan Parthasarathy,
Matthew Goldberg, Roger B. Dannenberg, Philippe Marchal, Ed Pervin.
{\it The Garnet Reference Manuals}.  Carnegie Mellon University Computer
Science Department Technical Report, no. CMU-CS-90-117-R5, Sep. 1994.
Revised from CMU-CS-90-117-R4, Oct. 1993, CMU-CS-90-117-R3, Nov. 1992,
CMU-CS-90-117-R2, May 1992,
CMU-CS-90-117-R, June 1991, CMU-CS-90-117, March, 1990, and
CMU-CS-89-196, Nov. 1989.

Bonnie E. John, Philip L. Miller, Brad A. Myers, Christine M.
Neuwirth, and Steven A. Shafer, eds. {\it Human-Computer Interaction in
the School of Computer Science}.  Carnegie Mellon University School of
Computer Science Technical Report, no. CMU-CS-92-193, October, 1992.

Brad A. Myers. {\it State of the Art in User Interface Software Tools}.
Carnegie Mellon University School of Computer Science Technical
Report, no. CMU-CS-92-114, February, 1992.

Brad A. Myers and Mary Beth Rosson.  {\it Survey on User Interface
Programming}.  Carnegie Mellon University School of Computer Science
Technical Report, no. CMU-CS-92-113, February, 1992.  Also published
as IBM Research Report RC17624.

Brad A. Myers, editor.  {\it The Garnet Compendium: Collected Papers,
1989-1990}.  Carnegie Mellon University School of Computer Science
Technical Report, no. CMU-CS-90-154, August, 1990.

David Kosbie, Brad Vander Zanden, Brad A. Myers, and Dario Giuse.
``Automatic Graphical Output Management,'' {\it The Garnet Compendium:
Collected Papers, 1989-1990}, Carnegie Mellon University School of
Computer Science Technical Report, no. CMU-CS-90-154, August, 1990,
pp. 29-43.

Roger B. Dannenberg, Brad A. Myers, Dario Giuse, and David Kosbie.
``Using Aggregates as Prototypes,'' {\it The Garnet Compendium:
Collected Papers, 1989-1990}, Carnegie Mellon University School of
Computer Science Technical Report, no. CMU-CS-90-154, August, 1990,
pp. 79-93.

Dario Giuse. {\it KR: Constraint-Based Knowledge Representation}.
Carnegie Mellon University Computer Science Department Technical Report,
no. CMU-CS-89-142, Apr, 1989.

Brad A. Myers. {\it The Garnet User Interface Development Environment; A
Proposal}.  Carnegie Mellon University Computer Science Department
Technical Report, no. CMU-CS-88-153, Sept, 1988.

Brad A. Myers. {\it Tools for Creating User Interfaces: An Introduction
and Survey,} Carnegie Mellon University Computer Science Department
Technical Report, no. CMU-CS-88-107, Jan, 1988.

Dario Giuse. {\it KR: an Efficient Knowledge Representation System}.
Carnegie Mellon University Robotics Institute Technical Report, no.
CMU-RI-TR-87-23, Oct, 1987.


%\vspace{1 line}
{\bf Submitted for Publication}

James Landay and Brad A. Myers.  ``Interactive Sketching for the Early
Stages of User Interface Design.''

Nobuhisa Yoda and Brad A. Myers.  ``A High Level Architecture with
Reusable Components for Synchronous Groupware Applications.''

Brad Vander Zanden and Brad A. Myers.  ``The Lapidary Graphical
Interface Design Tool.''



